\textcolor{blue}{Local Maximum Entropy Material Point Method applied to quasi-brittle fracture.}

### 2020-10-09 2020_EFM_MPM_Eigensoftening.tex
symbols

\usepackage{siunitx}
\usepackage[record,% using bib2gls
 symbols, % create list of symbols
 acronym
]{glossaries-extra}

\GlsXtrLoadResources[
  src={symbols}, % data in symbols.bib
  sort-field={name}, % sort by name field
  sort={letter-nocase}, % case-insensitive letter sort
  type=symbols, % put these terms in the symbols list
  field-aliases={unit=symbol}% convert unit key to symbol
]

@@

\newglossaryentry{domain}
{
  name=$\varOmega$,
  description={Continuum domain}
}

\newglossaryentry{contour}
{
  name=$\partial\varOmega$,
  description={Boundary of the continuum domain $\varOmega$. Also
    defined as $\Gamma$}
}

\newglossaryentry{dirichlet-boundary}
{
  name=$ \Gamma_d$,
  description={Essential or Dirichlet boundary conditions over $\partial\varOmega$}
}

\newglossaryentry{neumann-boundary}
{
  name=$ \Gamma_n$,
  description={Natural or Nemann boundary conditions over $\partial\varOmega$}
}

\newglossaryentry{a}
{
  name=$\vect{a}$,
  description={First order tensor which describes the acceleration field}
}

\newglossaryentry{v}
{
  name=$\vect{v}$,
  description={First order tensor which describes the velocity field}
}

\newglossaryentry{u}
{
  name=$\vect{u}$,
  description={First order tensor which describes the displacement field}
}

\newglossaryentry{x}
{
  name=$\vect{x}$,
  description={First order tensor which describes the global coordinates field}
}

\newglossaryentry{xi}
{
  name=$\vect{\xi}$,
  description={First order tensor which describes the local coordinates field}
}

\newglossaryentry{stress}
{
  name=$\tens{\sigma}$,
  description={Second order tensor which means the Cauchy stress tensor}
}

\newglossaryentry{strain}
{
  name=$\tens{\varepsilon}$,
  description={Second order tensor which means the Cauchy strain tensor}
}

\newglossaryentry{Constitutive}
{
  name=$\tens{D}$,
  description={Four order tensor which means the constitutive response
    of the material}
}

\newglossaryentry{LME-beta}
{
  name=$\beta$,
  description={Regularization or thermalization parameter of the
    LME$_{\beta}$ Pareto set}
}

\newglossaryentry{rho}
{
  name=$\rho$,
  description={Describes the scalar density field}
}

\newglossaryentry{fc}
{
  name=f$_c$,
  description={Compressive strenght}
}

\newglossaryentry{ft}
{
  name=f$_t$,
  description={Tensile strenght}
}

\newglossaryentry{Gf}
{
  name=G$_F$,
  description={Specific fracture energy}
}

\newglossaryentry{E}
{
  name=E,
  description={Elastic modulus}
}

\newglossaryentry{nu}
{
  name=$\nu$,
  description={Poisson ratio}
}

\newglossaryentry{Ceps}
{
  name=$C_{\epsilon}$,
  description={Normalizing constant}
}

\newglossaryentry{heps}
{
  name=$h_{\epsilon}$,
  description={Band width}
}

\newglossaryentry{Wc}
{
  name=$w_{c}$,
  description={Critical opening displacement}
}

\newglossaryentry{gamma-npc}
{
  name=$\gamma$,
  description={Time integration parameter for the \acrshort{npc} algorithm}
}

\newglossaryentry{gamma-lme}
{
  name=$\widehat{\gamma}$,
  description={Dimensionless parameter to control the value of the
    thermalization parameter $\beta$ of the \acrshort{lme} shape functions.}
}

@@

\textcolor{blue}{The \acrshort{mpm}~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian method where the continuum \gls{domain} is discretised with a finite set of material points (also known as particles in this research) $\hat{\Omega} = \{ \vec{x}_p, p \in \mathcal{C} \} \subset \Omega$, where $\mathcal{C} = 1, \ldots, n_p$. Any particle field such as position, velocity, mass, volume and stress denoted by $\vec{x}_p$, $\vec{v}_p$, $m_p$, $V_p$ and $\tens{\sigma}_p$, respectively, are assigned to each material point.  In addition to this set of material points, a set of fixed background nodes is introduced to compute the balance of momentum equation. This set of nodes is defined as  $\textbf{X} = \{ \vec{x}_I, I \in \mathcal{B} \} \subset \mathbb{R}^{2}$, where $\mathcal{B} = 1, \ldots, n_I$. In the present research, any nodal quantity is denoted by ($\square_I$).}

Since the \acrshort{mpm} possesses the advantages of both Lagrangian and Eulerian descriptions, no  element distortion takes place in the \acrshort{mpm}. Therefore, it is an appropriate and efficient method
to solve problems with moving discontinuities such as fracture evolution.\\
Without loosing generality, the \acrshort{mpm} algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where \textcolor{red}{the} balance of momentum equation is \textcolor{blue}{computed in the set of background nodes}, finally
(iii) a Lagrangian advection of the particles. In the present research,
\textcolor{blue}{we adopted the explicit predictor-corrector time integration scheme proposed by Molinos {\it et al.} \cite{Molinos2020}. The purpose of this choice is motivated due to its proven robustness and stability for dynamic computations, Navas {\it et al.} \cite{Navas2018a}}. In the first stage, the nodal velocity predictor \textcolor{red}{$ \vec{v}_{I}^{pred}$} is computed through
\textcolor{red}{particles} velocities \textcolor{red}{$\vec{v}_p$} and accelerations \textcolor{red}{$\vec{a}_p$} and evaluated
in the $k$ time step,
%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
 \label{eq:Predictor-velocity}
    \vec{v}_{I}^{pred} =  \textcolor{blue}{\sum_{p\ \in\ \mathcal{C}}} \frac{ N_{I}(\vec{x}_p^k )\ m_p\ (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%
Where \gls{gamma-npc} is an user-defined parameter (typically adopted as
0.5), $\tens{m}_I$ represents the lumped mass matrix, and \textcolor{red}{$N_{I}(\vec{x}_p^k) $} is the nodal contribution of the shape function evaluated in the particle
coordinates $\vec{x}_p^k$. This way of computing the nodal predictor is numerically stable
as well as minimises the computational effort. Once nodal velocities are
obtained, the essential boundary conditions are imposed. After that, an
Eulerian phase is computed in the set of nodes in a
\acrshort{fem}-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are
computed through the equilibrium equation. Next the nodal velocities
are corrected in a \textit{corrector} stage:
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\
  \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}\ .
\end{equation}
Finally, in order to update the particles, those are advected in the Lagrangian stage as
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} = \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}}
          \frac{N_{I}(\vec{x}_p^k)\ \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}} \Delta t\
        \frac{N_{I}(\vec{x}_p^k)\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n +  \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}} \Delta t\
         N_{I}(\vec{x}_p^k)\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{I}(\vec{x}_p^k)\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ .
\end{align}

@@

parameter\footnote{To avoid confusion with the \gls{gamma-npc} parameter of

@@

  \gls{gamma-lme} to preserve as much as possible the original notation.},

@@

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented \textcolor{red}{in this section} is to assess the
capability of the \acrshort{lme} approximants to improve the result \textit{versus} the standard linear interpolation and \acrshort{ugimp}
\cite{Bardenhagen2004}. The application
consists of a square plate of size $H = 1$ containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus \gls{E} = 1.06, the Poisson's ratio \gls{nu} = 0.333, and
critical energy-release rate \gls{Gf} = 0.0001. The absence of units is due to the current
simulation will not be tested against experimental results, where scale factor is relevant, but will be validated against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}. In \acrshort{mpm}, two different discretisation are required. On one
hand, a cartesian grid of nodes is considered with a nodal spacing value of 0.025. On the other hand, the plate will be modeled with a
initial layout of four particles per element and occupying the Gauss
quadrature positions.\\

@@

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test}. The experiment consist of loading
semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm, \textcolor{red}{and} supported by two rollers at a span of 47.5 mm, by
another roller on its top mid-span. Notches with a length of 16.6 mm are
created with the different notch angles of 0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$ with respect
to the vertical axis in order to investigate the influence of the
notch angle on the peak load \ref{fig:plot-Angle-Forces-test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material properties of Johnstone rock are listed in Table \ref{tab:Johnstone-properties}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering
  \begin{tabular}[]{l l}
    \hline
    Young's modulus (\gls{E})   & 0.4\  \textcolor{red}{GPa}       \\
    Poisson ratio (\gls{nu})    & 0.25           \\
    Density (\gls{rho})         & 1.54\ \textcolor{red}{$\text{g/cm}^3$} \\
    Tensile strength (\gls{ft}) & 0.6\ \textcolor{red}{MPa}       \\
    Critical opening displacement (\gls{Wc}) & 0.015\ \textcolor{red}{mm} \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the tests, particles were uniformly distributed with an almost
constant separation of 1 mm between them, resulting in a total number of 4672
\acrshort{mpm} particles for generating the specimen. Additionally, a
regular cartesian background grid of 1917 nodes with 1.5 mm of nodal distance
was generated. The velocity \textcolor{blue}{of the particles in the top boundary} is $v_y = 2 mm/s$, 
whereas the velocity of the bottom boundary particles is $v_y = 0 mm/s$. \textcolor{blue}{This boundary condition has been imposed via a soft loading ramp to ovoid dynamic effects due to a sudden loading}.
Finally, the model was calibrated with a regularization parameter \gls{Ceps} = 1.7 and a
bandwidth value \gls{heps} = 15.

@@

An interesting case of this experimental test was reported by Zhang {\it et al.} \cite{Zhang_2010a}. As in \textcolor{red}{this} study, an impact hammer of 120.6 kg has been employed in the proposed simulations to drop it at an impact
speed of 2640 mm/s. The beam dimensions were 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests (see Figure
\ref{fig:geometry-drop-weight-impact-test}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, being the material properties, such
as the material density, \gls{rho}, the compressive strenght, \gls{fc},
the tensile strenght \gls{ft}, the specific fracture energy, \gls{Gf},
and the elastic modulus \gls{E}, provided in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented in order to properly capture the impact. Several levels of
discretisation were required to assess the objectiveness of the
obtained results. The results here exposed are from a discretisation
of 21576 material points and 3943 nodes. To achieve better results, an
unstructured grid layout has been adopted focusing the minimum nodal
spacing, 1.2 mm, in the middle of the beam and in the impact surface
of the hammer. For this simulation, \gls{gamma-lme} was fixed to
1.8, this low value of the regularization parameter smooth high stress
overshots after the impact. Here the stabilization properties of the \acrshort{lme}
described in \cite{Molinos2020} becomes crucial to achieve
non-spurious results. Standard linear element were tested
with unsatisfactory results since grid-crossing occurs in the hammer
head producing non realistic forces.\\

@@

\printglossaries

>>>>

\usepackage[symbols,
  stylemods={tree},% loads glossaries-extra-stylemods to patch styles
  record % using 'bib2gls'
]{glossaries-extra}

% assign titles to group labels:
\glsxtrsetgrouptitle{latin}{Latin}
\glsxtrsetgrouptitle{greek}{Greek}

\GlsXtrLoadResources[
 src={symbols-latin},
 type=symbols,
 group={latin},% assign group label
 set-widest,% needed for 'alttree' styles
 save-locations=false
]

\GlsXtrLoadResources[
 src={symbols-greek},
 type=symbols,
 group={greek},% assign group label
 set-widest,% needed for 'alttree' styles
 save-locations=false
]

%\usepackage{siunitx}
%\usepackage[record,% using bib2gls
% acronym
%]{glossaries-extra}

%\GlsXtrLoadResources[
%  src={symbols}, % data in symbols.bib
%  sort-field={name}, % sort by name field
%  sort={letter-nocase}, % case-insensitive letter sort
%  type=symbols, % put these terms in the symbols list
%  field-aliases={unit=symbol}% convert unit key to symbol
%]

@@

%\newglossaryentry{domain}
%{
%  name=$\varOmega$,
%  description={Continuum domain}
%}

%\newglossaryentry{contour}
%{
%  name=$\partial\varOmega$,
%  description={Boundary of the continuum domain $\varOmega$. Also
%defined as $\Gamma$}
%}

%\newglossaryentry{dirichlet-boundary}
%{
%  name=$ \Gamma_d$,
%  description={Essential or Dirichlet boundary conditions over $\partial\varOmega$}
%}

%\newglossaryentry{neumann-boundary}
%{
%  name=$ \Gamma_n$,
%  description={Natural or Nemann boundary conditions over $\partial\varOmega$}
%}

%\newglossaryentry{a}
%{
%  name=$\vect{a}$,
%  description={First order tensor which describes the acceleration field}
%}

%\newglossaryentry{v}
%{
%  name=$\vect{v}$,
%  description={First order tensor which describes the velocity field}
%}

%\newglossaryentry{u}
%{
%  name=$\vect{u}$,
%  description={First order tensor which describes the displacement field}
%}

%\newglossaryentry{x}
%{
%  name=$\vect{x}$,
%  description={First order tensor which describes the global coordinates field}
%}

%\newglossaryentry{xi}
%{
%  name=$\vect{\xi}$,
%  description={First order tensor which describes the local coordinates field}
%}

%\newglossaryentry{stress}
%{
%  name=$\tens{\sigma}$,
%  description={Second order tensor which means the Cauchy stress tensor}
%}

%\newglossaryentry{strain}
%{
%  name=$\tens{\varepsilon}$,
%  description={Second order tensor which means the Cauchy strain tensor}
%}

%\newglossaryentry{Constitutive}
%{
%  name=$\tens{D}$,
%  description={Four order tensor which means the constitutive response
%    of the material}
%}

%\newglossaryentry{LME-beta}
%{
%  name=$\beta$,
%  description={Regularization or thermalization parameter of the
%    LME$_{\beta}$ Pareto set}
%}

%\newglossaryentry{rho}
%{
%  name=$\rho$,
%  description={Describes the scalar density field}
%}

%\newglossaryentry{fc}
%{
%  name=f$_c$,
%  description={Compressive strenght}
%}

%\newglossaryentry{ft}
%{
%  name=f$_t$,
%  description={Tensile strenght}
%}

%\newglossaryentry{Gf}
%{
%  name=G$_F$,
%  description={Specific fracture energy}
%}

%\newglossaryentry{E}
%{
%  name=E,
%  description={Elastic modulus}
%}

%\newglossaryentry{nu}
%{
%  name=$\nu$,
%  description={Poisson ratio}
%}

%\newglossaryentry{Ceps}
%{
%  name=$C_{\epsilon}$,
%  description={Normalizing constant}
%}

%\newglossaryentry{heps}
%{
%  name=$h_{\epsilon}$,
%  description={Band width}
%}

%\newglossaryentry{Wc}
%{
%  name=$w_{c}$,
%  description={Critical opening displacement}
%}

%\newglossaryentry{gamma-npc}
%{
%  name=$\gamma$,
%  description={Time integration parameter for the \acrshort{npc} algorithm}
%}

%\newglossaryentry{gamma-lme}
%{
%  name=$\widehat{\gamma}$,
%  description={Dimensionless parameter to control the value of the
%    thermalization parameter $\beta$ of the \acrshort{lme} shape functions.}
%}

@@

\textcolor{blue}{The \acrshort{mpm}~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian method where the continuum $\gls{domain}$ is discretised with a finite set of material points (also known as particles in this research) $\hat{\Omega} = \{ \vec{x}_p, p \in \mathcal{C} \} \subset \Omega$, where $\mathcal{C} = 1, \ldots, n_p$. Any particle field such as position, velocity, mass, volume and stress denoted by $\vec{x}_p$, $\vec{v}_p$, $m_p$, $V_p$ and $\tens{\sigma}_p$, respectively, are assigned to each material point.  In addition to this set of material points, a set of fixed background nodes is introduced to compute the balance of momentum equation. This set of nodes is defined as  $\textbf{X} = \{ \vec{x}_I, I \in \mathcal{B} \} \subset \mathbb{R}^{2}$, where $\mathcal{B} = 1, \ldots, n_I$. In the present research, any nodal quantity is denoted by ($\square_I$).}

Since the \acrshort{mpm} possesses the advantages of both Lagrangian and Eulerian descriptions, no  element distortion takes place in the \acrshort{mpm}. Therefore, it is an appropriate and efficient method
to solve problems with moving discontinuities such as fracture evolution.\\
Without loosing generality, the \acrshort{mpm} algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where \textcolor{red}{the} balance of momentum equation is \textcolor{blue}{computed in the set of background nodes}, finally
(iii) a Lagrangian advection of the particles. In the present research,
\textcolor{blue}{we adopted the explicit predictor-corrector time integration scheme proposed by Molinos {\it et al.} \cite{Molinos2020}. The purpose of this choice is motivated due to its proven robustness and stability for dynamic computations, Navas {\it et al.} \cite{Navas2018a}}. In the first stage, the nodal velocity predictor \textcolor{red}{$ \vec{v}_{I}^{pred}$} is computed through
\textcolor{red}{particles} velocities \textcolor{red}{$\vec{v}_p$} and accelerations \textcolor{red}{$\vec{a}_p$} and evaluated
in the $k$ time step,
%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
 \label{eq:Predictor-velocity}
    \vec{v}_{I}^{pred} =  \textcolor{blue}{\sum_{p\ \in\ \mathcal{C}}} \frac{ N_{I}(\vec{x}_p^k )\ m_p\ (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}.
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%
Where $\gls{gamma-npc}$ is an user-defined parameter (typically adopted as
0.5), $\tens{m}_I$ represents the lumped mass matrix, and \textcolor{red}{$N_{I}(\vec{x}_p^k) $} is the nodal contribution of the shape function evaluated in the particle
coordinates $\vec{x}_p^k$. This way of computing the nodal predictor is numerically stable
as well as minimises the computational effort. Once nodal velocities are
obtained, the essential boundary conditions are imposed. After that, an
Eulerian phase is computed in the set of nodes in a
\acrshort{fem}-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are
computed through the equilibrium equation. Next the nodal velocities
are corrected in a \textit{corrector} stage:
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\
  \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}\ .
\end{equation}
Finally, in order to update the particles, those are advected in the Lagrangian stage as
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} = \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}}
          \frac{N_{I}(\vec{x}_p^k)\ \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}} \Delta t\
        \frac{N_{I}(\vec{x}_p^k)\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ ,\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n +  \textcolor{blue}{\sum_{I\ \in\ \mathcal{B}_p}} \Delta t\
         N_{I}(\vec{x}_p^k)\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{I}(\vec{x}_p^k)\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\ .
\end{align}

@@

parameter\footnote{To avoid confusion with the $\gls{gamma-npc}$ parameter of

@@

  $\gls{gamma-lme}$ to preserve as much as possible the original notation.},

@@

\subsection{Edge-cracked square panel in mode I}
\label{sec:3.1}
The aim of the problem presented \textcolor{red}{in this section} is to assess the
capability of the \acrshort{lme} approximants to improve the result \textit{versus} the standard linear interpolation and \acrshort{ugimp}
\cite{Bardenhagen2004}. The application
consists of a square plate of size $H = 1$ containing an initial edge
crack of length $0.25 \cdot H$ loaded in a pure mode I by displacement
control on the outer flanks of the plate (See Figure
\ref{fig:geometry-cracked-panel-mode-I} for details). 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{Figure-Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The constitutive model considered in the numerical experiment is a
linear-elastic Hookean material, whose Young's modulus $\gls{E}$ = 1.06, the Poisson's ratio $\gls{nu}$ = 0.333, and
critical energy-release rate $\gls{Gf}$ = 0.0001. The absence of units is due to the current
simulation will not be tested against experimental results, where scale factor is relevant, but will be validated against an analytical
solution provided by Pandolfi \& Ortiz \cite{Pandolfi_2012}. In \acrshort{mpm}, two different discretisation are required. On one
hand, a cartesian grid of nodes is considered with a nodal spacing value of 0.025. On the other hand, the plate will be modeled with a
initial layout of four particles per element and occupying the Gauss
quadrature positions.\\

@@

The geometry and boundary conditions are sketched in Figure 
\ref{fig:geometry-Semicircular-bending-test}. The experiment consist of loading
semicircular specimens with a radius of 47.5 mm and a thickness of
20 mm, \textcolor{red}{and} supported by two rollers at a span of 47.5 mm, by
another roller on its top mid-span. Notches with a length of 16.6 mm are
created with the different notch angles of 0$^{\circ}$, 30$^{\circ}$ and 60$^{\circ}$ with respect
to the vertical axis in order to investigate the influence of the
notch angle on the peak load \ref{fig:plot-Angle-Forces-test}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.6\textwidth]{./Figure-Semicircular-bending-test}
  \caption{Geometry and boundary condition of the semicircular bending
    test.}
  \label{fig:geometry-Semicircular-bending-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material properties of Johnstone rock are listed in Table \ref{tab:Johnstone-properties}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
  \centering
  \begin{tabular}[]{l l}
    \hline
    Young's modulus ($\gls{E}$)   & 0.4\  \textcolor{red}{GPa}       \\
    Poisson ratio ($\gls{nu}$)    & 0.25           \\
    Density ($\gls{rho}$)         & 1.54\ \textcolor{red}{$\text{g/cm}^3$} \\
    Tensile strength ($\gls{ft}$) & 0.6\ \textcolor{red}{MPa}       \\
    Critical opening displacement ($\gls{Wc}$) & 0.015\ \textcolor{red}{mm} \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of Johnstone. ]{Material properties of Johnstone.}
  \label{tab:Johnstone-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In the tests, particles were uniformly distributed with an almost
constant separation of 1 mm between them, resulting in a total number of 4672
\acrshort{mpm} particles for generating the specimen. Additionally, a
regular cartesian background grid of 1917 nodes with 1.5 mm of nodal distance
was generated. The velocity \textcolor{blue}{of the particles in the top boundary} is $v_y = 2 mm/s$, 
whereas the velocity of the bottom boundary particles is $v_y = 0 mm/s$. \textcolor{blue}{This boundary condition has been imposed via a soft loading ramp to ovoid dynamic effects due to a sudden loading}.
Finally, the model was calibrated with a regularization parameter $\gls{Ceps}$ = 1.7 and a
bandwidth value $\gls{heps}$ = 15.

@@

An interesting case of this experimental test was reported by Zhang {\it et al.} \cite{Zhang_2010a}. As in \textcolor{red}{this} study, an impact hammer of 120.6 kg has been employed in the proposed simulations to drop it at an impact
speed of 2640 mm/s. The beam dimensions were 100 mm x 100 mm (B x D)
in cross section, and 420 mm in total lenght (L). The initial
notch-depth ratio was approximately 0.5, and the span, S, was fixed at
300 mm during the tests (see Figure
\ref{fig:geometry-drop-weight-impact-test}).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{./Figure-impact-test}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-drop-weight-impact-test}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The material adopted for the simulation was characterised by Navas
{\it et al.}~\cite{Navas_2017_ES}, being the material properties, such
as the material density, $\gls{rho}$, the compressive strenght, $\gls{fc}$,
the tensile strenght $\gls{ft}$, the specific fracture energy, $\gls{Gf}$,
and the elastic modulus $\gls{E}$, provided in table \ref{tab:concrete-properties}.\\
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
  \centering
  \begin{tabular}[]{c c c c c c c}
    \hline
      &   $\rho$   & $\text{f}_c$ & $\text{f}_t$ & $G_F$ &   E   & $\text{d}_{max}$ \\
      & (kg/m$^3$) &     (MPa)    &     (MPa)    & (N/m) & (GPa) & (mm) \\
    \hline
Value &    2368    &     102.7    &      5.4     &  141  &   31  &  12 \\
    \hline
  \end{tabular}
  \caption[Mechanical properties of thje concrete]{Mechanical
    properties of the high strenght concrete.}
  \label{tab:concrete-properties}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the present research, a 2D setting of the experiment has been
adopted. Both the hammer and the concrete beam are explicitly
represented in order to properly capture the impact. Several levels of
discretisation were required to assess the objectiveness of the
obtained results. The results here exposed are from a discretisation
of 21576 material points and 3943 nodes. To achieve better results, an
unstructured grid layout has been adopted focusing the minimum nodal
spacing, 1.2 mm, in the middle of the beam and in the impact surface
of the hammer. For this simulation, $\gls{gamma-lme}$ was fixed to
1.8, this low value of the regularization parameter smooth high stress
overshots after the impact. Here the stabilization properties of the \acrshort{lme}
described in \cite{Molinos2020} becomes crucial to achieve
non-spurious results. Standard linear element were tested
with unsatisfactory results since grid-crossing occurs in the hammer
head producing non realistic forces.\\

@@

%\printglossaries
\printunsrtglossary[type=symbols,style=alttreegroup,title={List of Symbols}]
